comgen-outline

### 2017-06-14 paper/comgen-paper.Rmd
finished draft of habituals results figure

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = F, fig.path='figs/',
                      echo=FALSE, warning=FALSE, cache=F, message=FALSE, sanitize = T)

@@

```{r habituals-prior-fig, fig.width = 8.5, fig.height = 4, fig.cap="Reconstructed priors on frequency from the structured, prior elicitation task (Expt. 2a). Priors are reconstructed by bootstrapping and forward sampling using the structured model (data, blue) or using Bayesian data analysis on the structured model (model, red)."}

@@

example.habituals <- c("climbs mountains_year", "hikes_year", "runs_year",
                       "goes to the movies_month", "smokes cigarettes_month")
fig.habituals.endorsement.vs.freq <- 
  ggplot(d.hab.endorsement.bayes, aes( x = logAnnualRate, y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_jitter(width = 0.1, shape = 21)+
  ylab("Human endorsement probability")+
  xlab("Frequency of target")+
  geom_label_repel(data = d.hab.endorsement.bayes %>% 
                    mutate(sentence = paste(habitual, time_period, sep = "_")) %>%
                    filter(sentence %in% example.habituals), aes( label = habitual ),
                       fontface = 'bold', color = 'white',
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.color = 'grey50')+
  scale_x_continuous(limits = c(-1,6), 
                     breaks = c(-1, 0, 0.7, 2.5, 4, 5.9),
                     labels = c("three years", "annually", "bi-annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-0.01, 1.01), 
                     breaks = c(0, 0.5, 1))+
  theme(axis.text.x = element_text(angle = 90))

@@

habituals.endorsement.models <- bind_rows(
  d.hab.endorse.regression %>% 
  mutate(prediction_lower = prediction,
         prediction_upper = prediction),
  left_join(
    m.hab.endorse.rsa %>% 
    rename(prediction = MAP, prediction_lower = cred_lower, prediction_upper = cred_upper),
    d.hab.endorse.regression %>% 
    select(habitual, time_period, logAnnualRate, mixture)
  )
) %>%
  mutate(src = factor(src, levels = c( 
                                      "habituals_model",
                                      "some_model",
                                      "regression_freq", 
                                      "regression_freq_distinct"
                                     ),
                      labels = c(
                                 "Speaker model - uncertain semantics",
                                 "Speaker model - fixed semantics",
                                 "Frequency alone", "Frequency + Distinctiveness"
                                 ))) %>%
  ggplot(., aes ( x = prediction, xmin = prediction_lower, xmax = prediction_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(shape = 21)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src, nrow = 2)+
  theme(legend.position = "bottom")

@@

```{r habituals-simulation-model, eval = F}
l0.hab.model <- '
var probability = function(Dist, x) {
    return Math.exp(Dist.score(x));
}
var targetUtterance = "habitual";

@@

var thetaBins = map2(function(b1, b2){
  var diff = Math.abs(b2 - b1) / 2;
  return roundTo3(diff+ b1);
}, stateBins(0, stateBins.length-1), stateBins.slice(1))

@@

  {
    "action": "climbs mountains",
    "mixture_female_mean": 0.0949,
    "mixture_female_samplesize": 3.7153,
    "mixture_male_mean": 0.0566,
    "mixture_male_samplesize": 0.9652,
    "stableFreq_female_mean": 0.3953,
    "stableFreq_female_samplesize": 1.6992,
    "stableFreq_male_mean": 0.8339,
    "stableFreq_male_samplesize": 1.5171
  }
]

var stable_female = beta({
  a: mixture_female_mean * mixture_female_samplesize,
  b: (1 - mixture_female_mean) * mixture_female_samplesize
})

var stable_male = beta({
  a: mixture_male_mean * mixture_male_samplesize,
  b: (1 - mixture_male_mean) * mixture_male_samplesize
})

var statePrior = Infer({model: function(){
  var gender = flip(0.5) ? "female" : "male";
  return gender == "female" ?
    flip(stable_female) ? 
      categorical({
        vs: stateBins,
        ps: map(function(b) {
          return probability(Gaussian({
                mu: stableFreq_female_mean, 
                sigma: stableFreq_female_samplesize
          }), b) + Number.EPSILON
        }, stateBins )
      }) :
    _.min(stateBins) : 
    flip(stable_male) ? 
      categorical({
        vs: stateBins,
        ps: map(function(b) {
          return probability(Gaussian({
                mu: stableFreq_male_mean, 
                sigma: stableFreq_male_samplesize
          }), b) + Number.EPSILON
        }, stateBins )
      }) :
    _.min(stateBins)
}});

//var out = {generic: listener0("generic"), some: listener0("some")}
//out
continuousListener0(data.utt[0])
'
```

```{r habituals-model-insets, fig.width = 4.75, fig.height = 1.5, eval = F}
example.habituals.actions <- separate(data.frame(example.habituals), 
                                      example.habituals, 
                                      into=c("action", "time_period"), sep = "_")$action

@@

 inputData = list(prior = list(params = data.frame(a = priorParams[["a"]],
                                                   b = priorParams[["b"]]),
                               mix = priorParams[["mix"]]), 
                  utt = "habitual")

@@

@@

```{r habituals-endorsement-figure, fig.width=11, fig.height=8}
grid.arrange(fig.habituals.endorsement.vs.freq, 
             habituals.endorsement.models,
             generics.endorsement.insets, ncol = 2,
             layout_matrix = cbind(c(1,1,3), c(2,2,3)))
```

>>>>

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = F, fig.path='figs/',
                      echo=FALSE, warning=FALSE, cache=T, message=FALSE, sanitize = T)

@@

```{r habituals-prior-fig, fig.width = 9, fig.height = 6, fig.cap="Reconstructed priors on frequency from the structured, prior elicitation task (Expt. 2a). Priors are reconstructed by bootstrapping and forward sampling using the structured model (data, blue) or using Bayesian data analysis on the structured model (model, red)."}

@@

example.habituals <- c("climbs mountains_year", "hikes_year", "runs_year",
                       "goes to the movies_month", "smokes cigarettes_month")
fig.habituals.endorsement.vs.freq <- 
  ggplot(d.hab.endorsement.bayes, aes( x = logAnnualRate, y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_jitter(width = 0.1, shape = 21)+
  ylab("Human endorsement probability")+
  xlab("Frequency of target")+
  geom_label_repel(data = d.hab.endorsement.bayes %>% 
                    mutate(sentence = paste(habitual, time_period, sep = "_")) %>%
                    filter(sentence %in% example.habituals), aes( label = habitual ),
                       fontface = 'bold', color = 'white',
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.color = 'grey50')+
  scale_x_continuous(limits = c(-1,6), 
                     breaks = c(-1, 0, 0.7, 2.5, 4, 5.9),
                     labels = c("three years", "annually", "bi-annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-0.01, 1.01), 
                     breaks = c(0, 0.5, 1))+
  theme(axis.text.x = element_text(angle = 90))+
  coord_fixed(ratio = 7)+
  guides(fill = F)
```

@@

habituals.endorsement.models <- bind_rows(
  d.hab.endorse.regression %>% 
  mutate(prediction_lower = prediction,
         prediction_upper = prediction),
  left_join(
    m.hab.endorse.rsa %>% 
    rename(prediction = MAP, prediction_lower = cred_lower, prediction_upper = cred_upper),
    d.hab.endorse.regression %>% 
    select(habitual, time_period, logAnnualRate, mixture)
  )
) %>%
  mutate(src = factor(src, levels = c( 
                                      "habituals_model",
                                      "some_model",
                                      "regression_freq", 
                                      "regression_freq_distinct"
                                     ),
                      labels = c(
                                 "Speaker model - uncertain semantics",
                                 "Speaker model - fixed semantics",
                                 "Frequency alone", "Frequency + Distinctiveness"
                                 ))) %>%
  ggplot(., aes ( x = prediction, xmin = prediction_lower, xmax = prediction_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(shape = 21)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src, nrow = 2)+
  theme(legend.position = "bottom")

@@

```{r habituals-simulation-model}
l0.hab.model <- '
var probability = function(Dist, x) {
    return Math.exp(Dist.score(x));
}
var betaShape = function(p){
  return {a: p.g * p.d, b: (1-p.g) * p.d}
};

var probBins = [0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9];

var targetUtterance = "habitual";

var roundTo3 = function(x){
  return Math.round(x * 10000) / 10000
}

@@

var thetaBins = map2(function(b1, b2){
  var diff = Math.abs(b2 - b1) / 2;
  return roundTo3(diff+ b1);
}, stateBins.slice(0, stateBins.length-1), stateBins.slice(1))

@@

var priorParams = data.priorParams[0];

var stable_female_params = betaShape({
  g: priorParams.mixture_female_mean, d: priorParams.mixture_female_samplesize
})

var stable_male_params = betaShape({
  g: priorParams.mixture_male_mean, d: priorParams.mixture_male_samplesize
})

// this marginalizes out the mixture parameter
var statePrior = Infer({model: function(){
  var gender = flip(0.5) ? "female" : "male";
  return gender == "female" ? // 
    flip(
        categorical({
          vs: probBins,
          ps: map(function(b) {
            return probability(Beta(stable_female_params), b) + Number.EPSILON
          }, probBins )
        })
    ) ? // stable female distribution
      categorical({
        vs: stateBins,
        ps: map(function(b) {
          return probability(Gaussian({
                mu: priorParams.stableFreq_female_mean, 
                sigma: priorParams.stableFreq_female_samplesize
          }), b) + Number.EPSILON
        }, stateBins )
      }) : // unstable female distribution
    _.min(stateBins) : 
    flip(        
      categorical({
          vs: probBins,
          ps: map(function(b) {
            return probability(Beta(stable_male_params), b) + Number.EPSILON
          }, probBins )
      })
    ) ? 
      categorical({
        vs: stateBins,
        ps: map(function(b) {
          return probability(Gaussian({
                mu: priorParams.stableFreq_male_mean, 
                sigma: priorParams.stableFreq_male_samplesize
          }), b) + Number.EPSILON
        }, stateBins )
      }) :
    _.min(stateBins)
}});

var listener0 = cache(function(utterance) {
  Infer({model: function(){
    var state = sample(statePrior)
    var state_prior = sample(statePrior)
    var theta = utterance == "habitual" ? sample(thetaPrior) : -99
    condition(meaning(utterance, state, theta))
    return {
      state_Posterior: state, 
      state_Prior: state_prior
  }
 }})}, 10000)

var continuousListener0 = function(utterance) {
  Infer({model: function(){
   var state = flip(0.5) ? 
        flip(beta(stable_female_params)) ? 
          gaussian({
                mu: priorParams.stableFreq_female_mean, 
                sigma: priorParams.stableFreq_female_samplesize
          }) : delta({v:-5}) :
        flip(beta(stable_male_params)) ? 
          gaussian({
                mu: priorParams.stableFreq_male_mean, 
                sigma: priorParams.stableFreq_male_samplesize
          }) : delta({v:-5})
   var state_prior = flip(0.5) ? 
        flip(beta(stable_female_params)) ? 
          gaussian({
                mu: priorParams.stableFreq_female_mean, 
                sigma: priorParams.stableFreq_female_samplesize
          }) : delta({v:lowerBound}) :
        flip(beta(stable_male_params)) ? 
          gaussian({
                mu: priorParams.stableFreq_male_mean, 
                sigma: priorParams.stableFreq_male_samplesize
          }) : delta({v:lowerBound})
    var theta = uniform(lowerBound, upperBound)
    condition(
      (utterance == "habitual") ? 
      state > theta : 
      (state > _.min(lowerBound))
    )
    return {
      state_Posterior: state, 
      state_Prior: state_prior
  }
 }, method: "rejection", samples: 10000, burn:5000, verbose: T})}

continuousListener0(data.utt[0])
'
```

```{r habituals-model-insets, fig.width = 4.75, fig.height = 1.5, cache = T}
example.habituals.actions <- separate(data.frame(example.habituals), 
                                      example.habituals, 
                                      into=c("action", "time_period"), sep = "_")$action

@@

 inputData = list(priorParams = priorParams,
                  utt = "habitual")

@@

# m.gen.fullmodel.target.prevalence <- m.samp %>%
#   filter(type == "withinKind") %>%
#   group_by(param, property, category) %>%
#   summarize(MAP = estimate_mode(val),
#             cred_upper = hdi_upper(val),
#             cred_lower = hdi_lower(val))

# m.gen.speakerBeliefs <- m.samp %>%
#   filter(type == "withinKind") %>%
#   mutate(Sentence = paste(category, property)) %>%
#   filter(Sentence %in% example.generics)

# m.gen.speakerBeliefs <- m.samp.prev.params <- m.samp %>%
#     filter(type == "targetPrevalence") %>%
#     mutate(Sentence = paste(category, property)) %>%
#     filter(Sentence %in% example.generics) %>%
#     mutate(parameter = paste(param, property, category, sep = "_")) %>%
#     select(-param, -property, -category, -chain, -type) %>%
#     group_by(parameter) %>%
#     mutate(iteration = ave(parameter==parameter, parameter, FUN=cumsum)) %>%
#     ungroup() %>%
#     separate(parameter, into = c("parameter", "property", "category"), sep= "_") %>%
#     group_by(category, property, iteration) %>%
#     spread(parameter, val) %>%
#     rowwise() %>%
#     mutate(
#       a = mean*sampleSize,
#       b = (1-mean)*sampleSize,
#       val = rbeta(n = 1, shape1 = a, shape2 = b)
#       ) %>%
#     ungroup()

# gen.inset.distributions <- bind_rows(
#   gen.listener.predictions %>%
#     mutate(category = NA),
#   m.gen.speakerBeliefs %>%
#     select(property, val) %>%
#     rename(value = val) %>%
#     mutate(Parameter = "speakerBeliefs")
# )
# 
# 
# category.text.labels <- data.frame(property = c("dont eat people", 
#                       "carry malaria", "lay eggs",  "are female", "have spots"),
#              category = c("Tigers", "Mosquitos", "Robins", "Robins", "Leopards"),
#              x = c(0.3, 0.3, 0.47, 0.05, 0.6),
#             y = c(0.45, 0.5, 0.45, 0.5, 0.26))

habituals.endorsement.insets <- hab.listener.predictions %>% 
    mutate(Parameter = factor(Parameter, levels = c("state_Prior",
                                                  "state_Posterior"
                                                  ),
                            labels = c("Listener Prior (Posterior given Silence)",
                                       "Listener Posterior given Habitual")),
         action = fct_relevel(action,
                                "climbs mountains", "hikes", "runs",
                              "goes to the movies", "smokes cigarettes"
                              ) ) %>%
  ggplot(., aes( x = value, fill = Parameter, color = Parameter, lty = Parameter, alpha = Parameter ))+
  geom_density(aes(y = ..scaled..), adjust = 4, size = 1)+
  facet_wrap(~action, nrow = 1)+
  # geom_text_repel(data = category.text.labels,
  #                 aes(label = category, x = x , y = y),
  #                 inherit.aes = F, color = "#2b83ba")+
  #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
  #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
  scale_fill_manual(values = c("#636363", "#d7191c"))+#, "#2b83ba"))+
  scale_color_manual(values = c("#636363", "#d7191c"))+#, "#2b83ba"))+
  scale_alpha_manual(values = c(0.6, 0.4))+#, 0))+
  #scale_linetype_manual(values = c(3, 4, 2, 1))+
  #scale_linetype_manual(values = c(3, 4, 1))+
  #scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
  scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
  xlab("Prevalence") +
  ylab("Scaled probability density")+
  theme(legend.position = "bottom", legend.title = element_blank(),
        axis.text.x = element_text(angle = 90))+
      scale_x_continuous(limits = c(-5, 8), 
                     breaks = c(-5, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))

@@

```{r habituals-endorsement-figure, fig.width=11, fig.height=8}
grid.arrange(fig.habituals.endorsement.vs.freq, 
             habituals.endorsement.models,
             habituals.endorsement.insets, ncol = 2,
             layout_matrix = cbind(c(1,1,3), c(2,2,3)))
```
